\thetitle

### 2020-07-01 writing/introduction.tex
first draft of introduction

Since its very beginning [cite STH], the theory of gradual typing
comes with a blame theorem. ``Well-typed [components]\footnote{The
original authors got this word wrong, and more.} can't go wrong'' has
become the slogan. The idea is that blame assignment can help the
practicing programmers find impedance mismatches between the types
they added and the remaining untyped components.

Practical gradual type systems ignore blame completely. Industrial
systems [lots of citations. footnotes] exploit types for assistance
with IDE actions and for finding typos in code. Then they remove them
and rely on the built-in safety checks of the underlying programming
language to catch problems.

This contrast raises the question whether

     blame adds value to gradual typing
     and especially the working programmer.

To obtain an answer, we have developed a method (not methodology!)  to
evaluate blame assignment systems from the gradual typing
world. Roughly speaking, the method simulates a programmer who follows
the above-cited Wadler-Findler slogan. When the run-time system
signals a violation, the programmer conjectures that the addition of 
types to untyped components turns the dynamic violation into a static
type error and thus reveals the impedance mismatch.

The paper presents the evaluation method and the results of applying
the method to three distinct blame assignment systems.

- explain the systems
- explain the method
- preview the results

- weave roadmap into these paragraphs

>>>>

From the very beginning, the theory of gradual typing has come with a blame
theorem~\cite{mf-toplas-2009, tf-dls-2006}. ``Well-typed [components]\footnote{The
original authors got this word wrong.} can't go wrong'' turned the theorem into a
slogan. Indeed, reasonably practical academic systems---for example,, Reticulated
Python~\cite{vsc-dls-2019, vss-popl-2017, vksb-dls-2014} and Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010}---come
with (different) blame assignment strategies. The idea is that blame assignment
can help the practicing programmers find impedance mismatches between the types
they added and the behavior remaining untyped components.

Practical gradual type systems ignore blame assignment almost completely.
Industrial systems such as Flow,\footnote{flow} Hack,\footnote{hack} or
TypeScript~\cite{ts} exploit types for assistance with IDE actions and for finding
typos in code. Their compilers remove types and rely on the built-in safety checks
of the underlying programming language to catch problems.

The contrast between academic and industrial systems raises the question whether
\begin{quote}
 \it
 blame assignment adds any value to a gradual typing system, \\
 especially for the working programmer.
\end{quote}
Given the prominence of the idea and the complete absence of blame in industrial
systems, it comes as an even bigger surprise that the research literature and the
industrial blog world do not present or discuss any possible answers.  Instead,
language designers seem to answer this question one way or another without any
scientific justification, and even when they have a language, they neglect to look
into this question. The research community simply does not have a method for
evaluating such questions, and given the question's practical relevance, this is a
problem.

This paper's first contribution is a method for evaluating the effectiveness of
blame assignment strategies in the gradual typing world (see
section~\ref{sec:ideas}). Roughly speaking, the method simulates a programmer who
follows the Wadler-Findler slogan quoted above. The implied method suggests that
programmers can find impedance-mismatch bugs by adding types to partially typed
software system. That is, when the run-time system signals a type-like violation,
a good blame assignment helps programmers infer where to add an additional type
specification. Since components that pass the type checker cannot cause run-time
exceptions; the type checker discovers these mismatches at compile time now, which
helps the programmer find mistake.  If the type specification implied by the blame
message does {\em not\/} cause a static type error, the program is run again until
it signals another violation.  Counting the number of steps it takes to discover
the mistake is a measure of the effectiveness of the blame assignment.  The method
is loosely based on the work of Lazarek et al. from last year's
POPL~\cite{lksfd-popl-2020}.

The paper's second contribution is a set of results from applying the evaluation
method to three distinct blame assignment systems and approximately 20,000
cases\mf{Replace with real number} (see section~\ref{sec:experiment}): Vitousek et
al.'s strategy of tracking typed/untyped boundaries in Reticulate; Tobin-Hochstadt
et al.'s use of Findler et al.'s higher-order contract system~\cite{ff-icfp-2002,
mf-toplas-2009}; and the error messages from failed safety checks in the
underlying programming language.

\mf{This is made up and must be fixed for the submission.}

The results of the experiment (see section~\ref{sec:results}) validate the
conjecture behind the work of theoreticians. A good blame assignment system
greatly shortens the search for impedance mismatches between the specified types
and the behavior of untyped components. Even better, the method clearly identifies
the wrapper-based blame tracking of Typed Racket as the vastly superior academic
system. Vitousek et al. lose one more time. Of course, a detailed discussion 
must acknowledge the threats to validity of these results (see
section~\ref{sec:discussion}).

The nest two sections provide an overview of the blame assignment strategies as
found in existing gradually typed language implementation (see
section~\ref{sec:landscape}) and the challenges of developing a systematic
evaluation method (section~\ref{sec:challengs}).
